Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-03-11 output/breed_paper_revised.Rnw
Adds non-breaking hypens.

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage{lscape} %% Harvard
%% \usepackage[shortcuts]{extdash}
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

%% \linenumbers

@@

Certainly one question in looking at this situation is, ``do new
cultivars have greater yield potential''?  Like many breeding
programs, one of the goals of the CCRRF is to increase yields through
breeding cultivars with increased yield potential. However, when the
popular cultivar M-202 was released in 1985, it was reported to yield
an average of 11 Mg ha\tsups{-1}~\citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M-209 was 10.8 Mg ha\tsups{-1}~\citep[14\%
MC;][]{m209}. However, in annual breeding reports and cultivar release
announcements, data based on yield performance in side-by-side cultivar
trials show that new cultivars usually out yield the older
cultivars, typically by 3--5\% (e.g., M-209 release description at \url{http://www.crrf.org/linked/2015annualreport.pdf}). In other rice systems, a decrease in a cultivar's yield
performance over time has been observed~\citep{Peng1998, Peng2010,
Peng2000, Datta1995}. It
has been speculated that this yield decline over time is due to the inability of
a cultivar to adapt to changing biotic and abiotic
conditions~\citep{Peng1998, Peng2010, Peng2000, Datta1995}. Yield
declines over time may explain the apparent contradiction above, in which
side-by-side yield comparisons show yield improvements in cultivars
while overall there may be little change in yield potential. To
address this issue we tested the hypothesis that if rice cultivar
yields erode over time then apparent yield increases in side-by-side
yield comparison tests will not reflect the true magnitude of yield
changes over time.

@@

Data were processed and models fit in R, an environment for
statistical computing~\citep{R}. Models were fit using the `rstanarm'
package~\citep{rstanarm}, an interface to Stan, a language for
probabilistic programming~\citep{stan}. All model diagnostics,
including $\hat{R}$, effective size, and posterior predictive checks
were examined before reporting results. To test goodness of fit, the
full model was compared to models containing only the years since
release or the year of release using the `loo' package~\citep{loo}, an
efficient means of conducting leave-one-out cross validation
\citep{loo_paper}. The 90\% credible interval (90\% CI), defined as
the interval containing 90\% of the distribution of estimated credible
parameter values given the data, was calculated as a measure of
uncertainty in the parameter estimates using the quantile method. The
90\% CI is more stable to sample to sample variance
and hence is preferable to the 95\% interval~\citep{rstanarm}. To
verify the estimates, the analysis was also conducted using classical
methods, specifically frequentist mixed-effects models fit by maximum
likelihood (see supplemental material for details and model
results). Complete data and code used for this analysis is available
at \url{github.com/mespe/rice_advances.git}.

@@

There was general agreement between the estimates from the Bayesian
and classical statistical approaches for all models, with the
estimates being similar between similar models. The Bayesian estimates
are presented here due to the inherent flexibility of Bayesian
inference, though the classical estimates can be found in the
supplemental material. Comparing the cultivars' yield during each's
initial time (average 3 years) in the experimental trials showed that across cultivars, yields ranged from
\Sexpr{round(plot_data["M207", "mean"], 2)} (M-207) to
\Sexpr{round(plot_data["M205", "mean"], 2)} Mg ha\tsups{-1} (M-205)
(Fig \ref{fig:first_yrs}). The relationship between the year
a cultivar was released and yield over the 33-year study period was
estimated to be slightly negative (\Sexpr{release_yr_coef}
kg\tsups{-1} ha\tsups{-1} yr) and does not provide
evidence of genetic yield improvements during this
period. Individual exceptions to this were two cultivars (M-204 and
M-205) that had higher yields than the others.

@@

Based on Model 3, there was no evidence of consistent differences between cultivars in
the rate of yield decline over time (i.e., all cultivar specific rates
of decline overlapped). Differences between cultivars in the decline
of yield over time, seen in the raw data (Fig. \ref{fig:all_cult}),
are therefore likely small and uncertain. Given this, the model excluding varying
rates of decline by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield decline over time and yield improvements over time (Model 2). The average decrease
in yield over time following a cultivar's release
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M-206 over M-202 (the current and former predominant
cultivars, respectively), can be attributed solely to the decline
of M-202's yield over time (Figs \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
decline, both M-202 and M-205 yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M-202 was
released 19 years prior to M-206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

@@

When yield decline over time is not accounted for in the analysis (Model 2),
there is estimated to be a yield increase of approximately
\Sexpr{only_release_coef} kg ha\tsups{-1} year\tsups{-1} (90\% CI
\Sexpr{only_release_CI[1]} to \Sexpr{only_release_CI[2]})
(Fig. \ref{fig:wo_erosion}). In contrast, the full model (Model 1) which
accounts for both the year of release and yield decline over time indicates that
the estimated change in grain yield per year in the CCRRF's breeding
trials at the RES was \Sexpr{release_yr_coef} kg ha\tsups{-1}
year\tsups{-1}, with both small positive and negative trends within
the credible interval (90\% CI: \Sexpr{rel_CI[1]} to
\Sexpr{rel_CI[2]}) (Fig. \ref{fig:first_yrs}). According to model
comparison metrics (PS-LOO, WAIC), Model 1 is
estimated to have superior predictive accuracy compared to the model
which does not account for yield decline over time.

After correcting for yield decline over time, some cultivars were higher
yielding compared to the overall average (as estimated via the random
intercept term in the model). The marginal estimate for the
performance of a cultivar averaged over the uncertainty in the other
parameters in the model (i.e., the random effect) shows three
cultivars with higher than average yields:
M-201, M-205, M-204 (Fig. \ref{fig:ids_comp}).

@@

@
\caption{The relationship between yield and the release year of medium
  grain rice cultivars for the last 6 years
  (\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by 
  correcting for yield decline over time. yield decline over time was corrected
  for by adjusting each observed yield by estimated amount lost 
  (\Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} after release).}
\label{fig:w_erosion}
\end{figure}
\end{comment}

@@

In CA, the cultivars with broad adaptation and adoption over the
33-year period of this study are M-202, M-206, and M-105. Cold
tolerance is a major problem for much of the California rice producing
region ~\citep{espe2017}, although it is a bigger problem in the
southern portion of the Sacramento Valley due to the influence of the
ocean winds through the Delta ~\citep{espe16a}. These three cultivars
have a higher degree of cold tolerance, high yield potential, and
relatively short duration and are thus broadly grown with over 70\% of
the area in these cultivars at any given time during the study
period~\citep{ucce}. There are several cultivars that were found to
have better than average yields, even after correcting for yield
declines over time. However, these cultivars (M-201, M-204, M-205, and possibly
M-209; Fig. \ref{fig:ids_comp}) are not widely adopted. Importantly,
not all medium grain cultivars that were released were purported to
have higher yields than previous cultivars. Some cultivars had special
traits, but upon release had similar yields to more broadly adapted
cultivars. Examples of this include M-208, which had resistance to
blast disease (\textit{Magnaporthe grisea}) and M-402 which is a
premium quality cultivar. Other cultivars such as M-205 and M-209,
while high yielding, do not have the cold tolerance of the other
cultivars and are thus only recommended for the northern portion of
the Sacramento Valley.

@@

The high yielding variety M-202 was released in 1985 and became the
most widely planted cultivar for two decades~\citep{mckenzie2014}. At
the time of release, M-202 yields were reported at 11.0 Mg
ha\tsups{-1}~\citep{m202} which is almost identical to what we . This
value of 11.0 Mg ha\tsups{-1} is what we also found in this analysis
based on the average experimental yields prior to public release
(Fig. \ref{fig:first_yrs}). Since the release of M-202, the rice
yields of newly released varieties cultivars have remained between
10.8 and 11.5 Mg ha\tsups{-1} (Fig. \ref{fig:first_yrs}), with the
exception of M-204 and M-205 (M-204 is no longer grown and M-205 has
limited adoaptation range so is only grown in a limited area). The
variety cultivar M-206 was released in 2004 and is currently still the
most widely grown rice varietycultivar in California. At the time of
its release, M-206 was reported to yield \Sexpr{round(9350*1.12/1000,1)} Mg
ha\tsups{-1} -- similar to what we found based on our analysis
(Fig. \ref{fig:first_yrs}). Based on this premise, it is clear that
improvements in yield potential have not been made over the 33-yr
period of this study. In contrast, both~\citet{Omar2014} and
~\citet{mckenzie2014}, report yield potential gains in California rice
systems based on analyses that showed that newly released cultivars
had higher yields than previously released cultivars. Based on their
analyses,~\citet{Omar2014} reported yield gains from the rice breeding
program of 37 to 60 kg ha\tsups{-1} yr\tsups{-1} and
~\citet{mckenzie2014} a yield increase of 25 kg ha\tsups{-1}
yr\tsups{-1}. The simpliest model (Model 2) in this study is similar
to the analyses of~\citet{Omar2014} and ~\citet{mckenzie2014}, and
results in similar findings; yields were estimated to increase by
\Sexpr{only_release_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{fig:wo_erosion}), almost identical to estimates by
~\citet{mckenzie2014}.

The only way to reconcile the two apparently contradictory findings above
is that the yield of potential of cultivars decline over
time such that when newly released cultivars are compared
with the older cultivar (as shown by ~\citet{Omar2014},
~\citet{mckenzie2014}, and Fig. \ref{fig:wo_erosion}), yields of the
new cultivar are higher. Here we provide evidence that
cultivar yield does decline - on average at a rate of
\Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{RES_year}). ~\citet{Peng2010} also found that a cultivar's
yield potential declines over time and reported a 15\%
decline over a 30 yr period.  In this study, over a 30 year period and
a \Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1} yield decline rate,
the yield decline would be approximately
\Sexpr{round((yrs_coef*-30)/mean(mod_data$yield_kg), 3)*100}\%. Given
the estimates here using the model which accounts for these declines
suggests that there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg\tsups{-1} ha\tsups{-1} year\tsups{-1}.

This analysis suggests that the CCRRF breeding program has
maintained but not substantially improved the yield potential in the
face of changing biotic and abiotic pressures over the 33 year time
frame of this study. This conclusion holds
true even when evaluating the cultivars which have been breed for
broad adaption and high yield potential (M-202, M-206, and
M-105).

@@

Yields of a particular cultivar may decline over time for a number of
reasons. First, ~\citet{Nie2009} reported yield declines in rice due
to continuously flooded conditions over a long time period which could
be overcome with improved N management. This is not likely the case
here as the fields used in this study grow rice only one season a year
and tend to be non-flooded for the remainder of the year. Furthermore,
at this location and unlike most CA rice fields, the fields are left
fallow every other year. Secondly, small changes in biotic and abiotic
forces have previously been speculated as the cause for yield decline
over time of the cultivar IR8 at IRRI~\citep{Peng1998, Peng2000,
  Peng2010}. Since the cultivar's genetics remain stable from the time
of selection but the conditions the cultivar experiences over time are
continuously changing, accumulated changes too small to reliably
measure could plausibly lead to yield declines over
time~\citep{Peng1998, Peng2000, Peng2010, Mackay2011}. However, it is
not clear if this mechanism is responsible for the yield decline over
time of \Sexpr{yrs_coef} Mg ha\tsups{-1} year\tsups{-1} found in this
analysis. Biotic stressors such as pests and diseases are low in this
Mediterranean climate. Furthermore, these fields are fallowed every
other year and the residues are burned, further lowering the potential
for pest and disease pressure. Thirdly, changing climate may influence
yield potential with either warmer~\citep{van2013} or
cooler~\citep{espe2017} temperatures reducing
yields. Yield potential can be defined as the yield of optimum
cultivar in the absence of biotic and abiotic stresses in a given
environment \citep{vanIttersum2013}. However, ~\citet{espe2017} found
no substantial trends in growing season temperatures (either high,
low, or average temps) in the Sacramento Valley during the study
period, suggesting that older cultivars do not have lower yield
potential due to changing climatic factors. Our analysis
offers inconclusive evidence of an interaction between cultivar and
year (i.e., the CI for the interactions overlapped with each other and zero), visual examination suggests that cultivars may be
different in terms of rates of yield decline
(Fig. \ref{fig:all_cult}); however, we do not have sufficient data nor
the proper controls in the existing data set to fully test this
hypothesis. More data is needed to further explore the exact cause of
the yield declines observed here, as well to determine if different
cultivars have differential declines over time.

@@

California state average yields have increased slowly during the study
period (Fig. \ref{fig:state_yield}) \citep{espe16a}. What are the
potential causes?  It is not due to the adoption of the broadly
adapted cultivars (M-206 and M-105) as there is no evidence of
increased yield potential among these (Fig. \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}) compared to M-202 released at the start of this
period. Although M-205 is not broadly adapted, it has
increase yield potential (Fig. \ref{fig:first_yrs} and its adoption (although limited) in the
northern and warmer part of CA may contribute to some yield gains
reported at the state level. Other factors include improvement in
agronomic management, including more efficient nutrient
management~\citep{Lundy2012, Lundy2015, linquist2009}, and weed
management~\citep{deforest2017, Pittelkow2012, caton2002}. However,
recent rates of state-wide yield gains are far less than those of the
preceding era, consistent with a system near the yield
ceiling~\citep{Grassini2013}.

\begin{comment}
\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg\tsups{-1} ha\tsups{-1}
year\tsups{-1}. Yet the CI estimated from these data using the best
model suggests that there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg\tsups{-1} ha\tsups{-1} year\tsups{-1}. There are several
explanations for this discrepancy in conclusions. First, \citet{mckenzie2014} did not
account for yield decreases over time in their analysis. As shown
here, when these decreases are not taken into about, the estimated
increase in yield over time is estimated to be approximately the same
as that found by \citet{mckenzie2014}. Second, these yield decreases over time
were not taken into account possibly because \citet{mckenzie2014}
concluded that there was no trend in the yield of the major cultivars
over time. However this conclusion is based on a flawed interpretation
of the results of their statistical test. Failing to reject the
hypothesis that the trend in yield equals zero does not equate to the
confirmation of no trend. Instead, failing to reject the hypothesis
could suggest the analysis lacks the statistical power to detect the
trend. Third, it
is not surprising that no trend in yield over time was detected given
that \citet{mckenzie2014} used data from several locations, yet
analyzed those data using a simple t-test. By using this simple
analysis, \citet{mckenzie2014} ignore systematic errors related to
differences between observations from different locations, including
different management systems, growing environments, and crop
timing. By contrast, the investigation presented here used a single,
well managed system and analyzed the data using a mixed effects model
which fully accounts for non-independence between similar
observations. In doing this, the analysis makes fewer assumptions and
more fully captures the structural relationships in the data. Hence,
the estimates presented here are likely more reliable.
\end{comment}

@@

@
\caption{The performance of two major cultivars over the period
\Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
solid line is the estimated yield decline over time (\Sexpr{yrs_coef} kg
ha\tsups{-1} year\tsups{-1}), while the dashed line is the published
3-year average yield of M-202 at its release (1982-1984). Both
cultivars have a similar yield at their respective year of release,
and similar decreases over time.}
\label{fig:compare_yrs2}
\end{figure}
\newpage

>>>>

\documentclass[12pt,]{elsarticle}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable,booktabs}
\usepackage{comment}
\usepackage{graphicx,grffile, textcomp}
\usepackage{setspace}
\usepackage{etoolbox}
\usepackage{float}
\doublespacing
\usepackage{lineno}
\usepackage[shortcuts]{extdash} %% Avoid breaking on M--20X
\usepackage{lscape} %% Harvard
%% \usepackage[shortcuts]{extdash}
\bibliographystyle{model2-names}\biboptions{authoryear}
\pagenumbering{gobble}
%% For figure numbers in supplementary info
%% from http://bytesizebio.net/2013/03/11/adding-supplementary-tables-and-figures-in-latex/
\newcommand{\beginsupplement}{%
  \setcounter{table}{0}
  \renewcommand{\thetable}{S\arabic{table}}%
  \setcounter{figure}{0}
  \renewcommand{\thefigure}{S\arabic{figure}}%
}

@@

\linenumbers

@@

Certainly one question in looking at this situation is, ``do new
cultivars have greater yield potential''?  Like many breeding
programs, one of the goals of the CCRRF is to increase yields through
breeding cultivars with increased yield potential. However, when the
popular cultivar M\=/202 was released in 1985, it was reported to yield
an average of 11 Mg ha\tsups{-1}~\citep[14\% moisture content
(MC);][]{m202}. In 2015, the reported average yield of the newest
released cultivar M\=/209 was 10.8 Mg ha\tsups{-1}~\citep[14\%
MC;][]{m209}. However, in annual breeding reports and cultivar release
announcements, data based on yield performance in side-by-side cultivar
trials show that new cultivars usually out yield the older
cultivars, typically by 3--5\% (e.g., M\=/209 release description at \url{http://www.crrf.org/linked/2015annualreport.pdf}). In other rice systems, a decrease in a cultivar's yield
performance over time has been observed~\citep{Peng1998, Peng2010,
Peng2000, Datta1995}. It
has been speculated that this yield decline over time is due to the inability of
a cultivar to adapt to changing biotic and abiotic
conditions~\citep{Peng1998, Peng2010, Peng2000, Datta1995}. Yield
declines over time may explain the apparent contradiction above, in which
side-by-side yield comparisons show yield improvements in cultivars
while overall there may be little change in yield potential. To
address this issue we tested the hypothesis that if rice cultivar
yields erode over time then apparent yield increases in side-by-side
yield comparison tests will not reflect the true magnitude of yield
changes over time.

@@

Data were processed and models fit in R, an environment for
statistical computing~\citep{R}. Models were fit using the `rstanarm'
package~\citep{rstanarm}, an interface to Stan, a language for
probabilistic programming~\citep{stan}. All model diagnostics,
including $\hat{R}$, effective size, and posterior predictive checks
were examined before reporting results. To test goodness of fit, the
full model was compared to models containing only the years since
release or the year of release using the `loo' package~\citep{loo}, an
efficient means of conducting leave-one-out cross validation
\citep{loo_paper}. The 90\% credible interval (90\% CI), defined as
the interval containing 90\% of the distribution of estimated credible
parameter values given the data, was calculated as a measure of
uncertainty in the parameter estimates using the quantile method. The
90\% CI is more stable to sample to sample variance
and hence is preferable to the 95\% interval~\citep{rstanarm}. To
verify the estimates, the analysis was also conducted using classical
methods, specifically frequentist mixed-effects models fit by maximum
likelihood (see supplemental material for details and model
results). Complete data and code used for this analysis is available
through the Open Science Framework at \url{https://osf.io/6ed5k/}.

@@

There was general agreement between the estimates from the Bayesian
and classical statistical approaches for all models, with the
estimates being similar between similar models. The Bayesian estimates
are presented here due to the inherent flexibility of Bayesian
inference, though the classical estimates can be found in the
supplemental material. Comparing the cultivars' yield during each's
initial time (average 3 years) in the experimental trials showed that across cultivars, yields ranged from
\Sexpr{round(plot_data["M207", "mean"], 2)} (M\=/207) to
\Sexpr{round(plot_data["M205", "mean"], 2)} Mg ha\tsups{-1} (M\=/205)
(Fig \ref{fig:first_yrs}). The relationship between the year
a cultivar was released and yield over the 33-year study period was
estimated to be slightly negative (\Sexpr{release_yr_coef}
kg ha\tsups{-1} yr\tsups{-1}) and does not provide
evidence of genetic yield improvements during this
period. Individual exceptions to this were two cultivars (M\=/204 and
M\=/205) that had higher yields than the others.

@@

Based on Model 3, there was no evidence of consistent differences between cultivars in
the rate of yield decline over time (i.e., all cultivar specific rates
of decline overlapped). Differences between cultivars in the decline
of yield over time, seen in the raw data (Fig. \ref{fig:all_cult}),
are therefore likely small and uncertain. Given this, the model excluding varying
rates of decline by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield decline over time and yield improvements over time (Model 2). The average decrease
in yield over time following a cultivar's release
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M\=/206 over M\=/202 (the current and former predominant
cultivars, respectively), can be attributed solely to the decline
of M\=/202's yield over time (Figs \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
decline, both M\=/202 and M\=/205 yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M\=/202 was
released 19 years prior to M\=/206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

@@

When yield decline over time is not accounted for in the analysis (Model 2),
there is estimated to be a yield increase of approximately
\Sexpr{only_release_coef} kg ha\tsups{-1} yr\tsups{-1} (90\% CI
\Sexpr{only_release_CI[1]} to \Sexpr{only_release_CI[2]})
(Fig. \ref{fig:wo_erosion}). In contrast, the full model (Model 1) which
accounts for both the year of release and yield decline over time indicates that
the estimated change in grain yield per year in the CCRRF's breeding
trials at the RES was \Sexpr{release_yr_coef} kg ha\tsups{-1}
yr\tsups{-1}, with both small positive and negative trends within
the credible interval (90\% CI: \Sexpr{rel_CI[1]} to
\Sexpr{rel_CI[2]}) (Fig. \ref{fig:first_yrs}). According to model
comparison metrics (PS-LOO, WAIC), Model 1 is
estimated to have superior predictive accuracy compared to the model
which does not account for yield decline over time.

After correcting for yield decline over time, some cultivars were higher
yielding compared to the overall average (as estimated via the random
intercept term in the model). The marginal estimate for the
performance of a cultivar averaged over the uncertainty in the other
parameters in the model (i.e., the random effect) shows three
cultivars with higher than average yields:
M\=/201, M\=/205, M\=/204 (Fig. \ref{fig:ids_comp}).

@@

@
\caption{The relationship between yield and the release year of medium
  grain rice cultivars for the last 6 years
  (\Sexpr{min(comp_yrs)} -- \Sexpr{max(comp_yrs)}) as estimated by 
  correcting for yield decline over time. yield decline over time was corrected
  for by adjusting each observed yield by estimated amount lost 
  (\Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1} after release).}
\label{fig:w_erosion}
\end{figure}
\end{comment}

@@

In CA, the cultivars with broad adaptation and adoption over the
33-year period of this study are M\=/202, M\=/206, and M\=/105. Cold
tolerance is a major problem for much of the California rice producing
region ~\citep{espe2017}, although it is a bigger problem in the
southern portion of the Sacramento Valley due to the influence of the
ocean winds through the Delta ~\citep{espe16a}. These three cultivars
have a higher degree of cold tolerance, high yield potential, and
relatively short duration and are thus broadly grown with over 70\% of
the area in these cultivars at any given time during the study
period~\citep{ucce}. There are several cultivars that were found to
have better than average yields, even after correcting for yield
declines over time. However, these cultivars (M\=/201, M\=/204, M\=/205, and possibly
M\=/209; Fig. \ref{fig:ids_comp}) are not widely adopted. Importantly,
not all medium grain cultivars that were released were purported to
have higher yields than previous cultivars. Some cultivars had special
traits, but upon release had similar yields to more broadly adapted
cultivars. Examples of this include M\=/208, which had resistance to
blast disease (\textit{Magnaporthe grisea}) and M\=/402 which is a
premium quality cultivar. Other cultivars such as M\=/205 and M\=/209,
while high yielding, do not have the cold tolerance of the other
cultivars and are thus only recommended for the northern portion of
the Sacramento Valley.

@@

The high yielding variety M\=/202 was released in 1985 and became the
most widely planted cultivar for two decades~\citep{mckenzie2014}. At
the time of release, M\=/202 yields were reported at 11.0 Mg
ha\tsups{-1}~\citep{m202} which is almost identical to what we . This
value of 11.0 Mg ha\tsups{-1} is what we also found in this analysis
based on the average experimental yields prior to public release
(Fig. \ref{fig:first_yrs}). Since the release of M\=/202, the rice
yields of newly released varieties cultivars have remained between
10.8 and 11.5 Mg ha\tsups{-1} (Fig. \ref{fig:first_yrs}), with the
exception of M\=/204 and M\=/205 (M\=/204 is no longer grown and M\=/205 has
limited adoaptation range so is only grown in a limited area). The
variety cultivar M\=/206 was released in 2004 and is currently still the
most widely grown rice varietycultivar in California. At the time of
its release, M\=/206 was reported to yield \Sexpr{round(9350*1.12/1000,1)} Mg
ha\tsups{-1} -- similar to what we found based on our analysis
(Fig. \ref{fig:first_yrs}). Based on this premise, it is clear that
improvements in yield potential have not been made over the 33-year
period of this study. In contrast, both~\citet{Omar2014} and
~\citet{mckenzie2014}, report yield potential gains in California rice
systems based on analyses that showed that newly released cultivars
had higher yields than previously released cultivars. Based on their
analyses,~\citet{Omar2014} reported yield gains from the rice breeding
program of 37 to 60 kg ha\tsups{-1} yr\tsups{-1} and
~\citet{mckenzie2014} a yield increase of 25 kg ha\tsups{-1}
yr\tsups{-1}. The simpliest model (Model 2) in this study is similar
to the analyses of~\citet{Omar2014} and ~\citet{mckenzie2014}, and
results in similar findings; yields were estimated to increase by
\Sexpr{only_release_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{fig:wo_erosion}), almost identical to estimates by
~\citet{mckenzie2014}.

The only way to reconcile the two apparently contradictory findings
above is that the yield of potential of cultivars decline over time
such that when newly released cultivars are compared with the older
cultivar (as shown by ~\citet{Omar2014}, ~\citet{mckenzie2014}, and
Fig. \ref{fig:wo_erosion}), yields of the new cultivar are
higher. Here we provide evidence that cultivar yield does decline - on
average at a rate of \Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{fig:RES_year}). When these declines are taken into account,
there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of a rate of yield increase equal to or greater than 25
kg ha\tsups{-1} yr\tsups{-1} during the study
period. ~\citet{Peng2010} also found that a cultivar's yield potential
declines over time and reported a 15\% decline over a 30 yr period.
In this study, over a 30 year period and a \Sexpr{yrs_coef*-1} kg
ha\tsups{-1} yr\tsups{-1} rate, the yield decline would be
approximately \Sexpr{round((yrs_coef*-30)/mean(mod_data$yield_kg),2)*100}\%.

This analysis suggests that the CCRRF breeding program has
maintained but not substantially improved the yield potential in the
face of changing biotic and abiotic pressures over the 33 year time
frame of this study. This conclusion holds
true even when evaluating the cultivars which have been breed for
broad adaption and high yield potential (M\=/202, M\=/206, and
M\=/105).

@@

Yields of a particular cultivar may decline over time for a number of
reasons. First, ~\citet{Nie2009} reported yield declines in rice due
to continuously flooded conditions over a long time period which could
be overcome with improved N management. This is not likely the case
here as the fields used in this study grow rice only one season a year
and tend to be non-flooded for the remainder of the year. Furthermore,
at this location and unlike most CA rice fields, the fields are left
fallow every other year. Secondly, small changes in biotic and abiotic
forces have previously been speculated as the cause for yield decline
over time of the cultivar IR8 at IRRI~\citep{Peng1998, Peng2000,
  Peng2010}. Since the cultivar's genetics remain stable from the time
of selection but the conditions the cultivar experiences over time are
continuously changing, accumulated changes too small to reliably
measure could plausibly lead to yield declines over
time~\citep{Peng1998, Peng2000, Peng2010, Mackay2011}. However, it is
not clear if this mechanism is responsible for the yield decline over
time of \Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1} found in this
analysis. Biotic stressors such as pests and diseases are low in this
Mediterranean climate. Furthermore, these fields are fallowed every
other year and the residues are burned, further lowering the potential
for pest and disease pressure. Thirdly, changing climate may influence
yield potential with either warmer~\citep{van2013} or
cooler~\citep{espe2017} temperatures reducing
yields. Yield potential can be defined as the yield of optimum
cultivar in the absence of biotic and abiotic stresses in a given
environment \citep{vanIttersum2013}. However, ~\citet{espe2017} found
no substantial trends in growing season temperatures (either high,
low, or average temps) in the Sacramento Valley during the study
period, suggesting that older cultivars do not have lower yield
potential due to changing climatic factors. While our analysis offers
inconclusive evidence of differences between cultivars in rate of
yield decline (i.e., the CIs for the interactions overlapped with
each other and zero), visual examination suggests that cultivars may
be different in terms of rates of yield decline
(Fig. \ref{fig:all_cult}); however, we do not have sufficient data nor
the proper controls in the existing data set to fully test this
hypothesis. More data is needed to further explore the exact cause of
the yield declines observed here, as well to determine if different
cultivars have differential declines over time.

@@

California state average yields have increased slowly during the study
period (Fig. \ref{fig:state_yield}) \citep{espe16a}. What are the
potential causes?  It is not due to the adoption of the broadly
adapted cultivars (M\=/206 and M\=/105) as there is no evidence of
increased yield potential among these (Fig. \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}) compared to M\=/202 released at the start of this
period. Although M\=/205 is not broadly adapted, it has
increase yield potential (Fig. \ref{fig:first_yrs}) and its adoption (although limited) in the
northern and warmer part of CA may contribute to some yield gains
reported at the state level. Other factors include improvement in
agronomic management, including more efficient nutrient
management~\citep{Lundy2012, Lundy2015, linquist2009}, and weed
management~\citep{deforest2017, Pittelkow2012, caton2002}. However,
recent rates of state-wide yield gains are far less than those of the
preceding era, consistent with a system near the yield
ceiling~\citep{Grassini2013}.

\begin{comment}
\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg ha\tsups{-1}
yr\tsups{-1}. Yet the CI estimated from these data using the best
model suggests that there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg ha\tsups{-1} yr\tsups{-1}. There are several
explanations for this discrepancy in conclusions. First, \citet{mckenzie2014} did not
account for yield decreases over time in their analysis. As shown
here, when these decreases are not taken into about, the estimated
increase in yield over time is estimated to be approximately the same
as that found by \citet{mckenzie2014}. Second, these yield decreases over time
were not taken into account possibly because \citet{mckenzie2014}
concluded that there was no trend in the yield of the major cultivars
over time. However this conclusion is based on a flawed interpretation
of the results of their statistical test. Failing to reject the
hypothesis that the trend in yield equals zero does not equate to the
confirmation of no trend. Instead, failing to reject the hypothesis
could suggest the analysis lacks the statistical power to detect the
trend. Third, it
is not surprising that no trend in yield over time was detected given
that \citet{mckenzie2014} used data from several locations, yet
analyzed those data using a simple t-test. By using this simple
analysis, \citet{mckenzie2014} ignore systematic errors related to
differences between observations from different locations, including
different management systems, growing environments, and crop
timing. By contrast, the investigation presented here used a single,
well managed system and analyzed the data using a mixed effects model
which fully accounts for non-independence between similar
observations. In doing this, the analysis makes fewer assumptions and
more fully captures the structural relationships in the data. Hence,
the estimates presented here are likely more reliable.
\end{comment}

@@

@
\caption{The performance of two major cultivars over the period
\Sexpr{min(mod_data[,'year'])} -- \Sexpr{max(mod_data[,'year'])}. The
solid line is the estimated yield decline over time (\Sexpr{yrs_coef} kg
ha\tsups{-1} yr\tsups{-1}), while the dashed line is the published
3-year average yield of M\=/202 at its release (1982-1984). Both
cultivars have a similar yield at their respective year of release,
and similar decreases over time.}
\label{fig:compare_yrs2}
\end{figure}
\newpage
